\section{Preliminaries}

The original Delta Correlating Prediction Tables algorithm can be seen in
algorithm \ref{alg:dcpt_orig}, as described in Grannæs et.
al.\cite{Grannaes481837}. Initially, the PC is looked up in a table of entries,
and if it is not found, a replacement entry is created. If it was in the table,
the new delta from the last address to the current is added to the list of
deltas. 

The tiered version of the DCPT exchanges the table of entries with different
tier tables, which are searched through sequentially. The TDCPT did not yield
any significant speedup, but did lower the memory footprint.

\begin{algorithm}
\footnotesize
\caption{Original DCPT \cite{Grannaes481837}}
  \begin{algorithmic}[1]
    \Procedure{DCPT}{Load address $PC$, Miss address $addr$}
      \State $r \gets \Call{TableLookUp}{PC}$ \\

      \If {$entry.pc \neq PC$}
        \State $entry.pc \gets PC$
        \State $entry.lastAddress \gets addr$
        \State $entry.deltas[\;] \gets \emptyset$
        \State $entry.lastPrefetch \gets 0$
      \ElsIf {$addr - entry.lastAddress \neq 0$}
        \State $entry.deltas[\;] \gets addr - entry.lastAddress$
        \State $entry.lastAddress \gets addr$ \\

        \State $candidates[\;] \gets \Call{DeltaCorrelation}{entry}$
        \State $prefetches[\;] \gets \Call{PrefetchFilter}{entry, candidates[\:]}$
        \State $\Call{IssuePrefetches}{prefetches[\:]}$
      \EndIf
    \EndProcedure
  \end{algorithmic}
  \label{alg:dcpt_orig}
\end{algorithm}

\begin{algorithm}
\footnotesize
\caption{Dynamic Tiered DCPT}
  \begin{algorithmic}[1]
    \Procedure{DCPT}{Load address $PC$, Miss address $addr$}
      \State $entry \gets \Call{TableLookUp}{PC}$

      \If {$entry.pc \neq PC$}
        \If {$mode = Tiered$}
          \State $entry \gets \Call{T1TableLookUp}{PC}$
          \If {$entry.pc = PC$}
            \State $\Call{MoveToT3}{entry}$
          \Else
	    \State $\Call{AddToT1}{entry}$ 
          \EndIf
        \Else
          \If {$t1\_hits\_ratio \geq tolerance$}
            \State $\Call{SwitchMode}{TIERED}$ \\
	    \State $\Call{AddToT1}{entry}$
          \Else
	    \State $\Call{AddToT3}{entry}$
          \EndIf
        \EndIf
      \ElsIf {$addr - entry.lastAddress \neq 0$}
	\If {$mode = Tiered \ \& $\\
	  \hspace{11mm}$t1\_hits\_ratio \leq tolerance - deadzone$}
	  \State $\Call{SwitchMode}{TIER3\_ONLY}$
	\EndIf
	\State $\Call{AddToT3}{entry}$
      \EndIf
    \EndProcedure
  \end{algorithmic}
  \label{alg:dcpt_dyn_tier}
\end{algorithm}


\subsection{Prefetcher Description}

Our prefetcher, Dynamically Tiered Delta Correlating Prediction
Tables (algorithm \ref{alg:dcpt_dyn_tier}) aims to have dynamically sized tiers to improve upon the potential
pitfalls that the original TDCPT-prefetcher fell into. Additionally it will
gather some statistics about the hit-rates for the various tiers to attempt to
decide better what size the various tiers should be, and whether or not to tier
at all.

A tier is here understood in the same way as in Sømåen et al, that is, when a
new PC is encountered, it doesn't immediately take up the space of a full-size
table entry, but is instead placed in a table of new PCs, along with the
address it accessed, this table is hereafter referred to as Tier 1 (T1). While
Sømåen et al. introduced a second tier (T2) for PCs that had been seen more
than once, but hadn't yet reached a point where they would have been seen
enough times to generate the necessary amount of deltas to perform any
prefetching, where they would be placed in Tier 3 (T3), we opted to skip this
tier to keep our solution a bit simpler with regards to the tiering structure.
To keep the language consistent, the tiers are still referred to using the
same terms as before, that is where T1 refers to the short table, and T3
refers to full length entries.

To compensate for the varying usage patterns in the programs, the prefetcher
uses the same table-memory in two different ways, either it uses it as a pure
T3-only setup, which would behave exactly like a pure DCPT solution, or it
uses a small portion of this to make room for a comparably large amount of
T1-elements. This should in theory help to buffer against a sudden large
influx of T1-type PCs, and thus avoids removal of T3 entries
prematurely.

To allow the parts of the programs that need more T3 elements than what is
available when using some of the table for T1, the prefetcher will switch
between the tiered operation mode , and pure DCPT as follows:

As an access happens, the algorithm checks if the PC accessed is in T3. If it's
not, and the state is tiered mode, the PC will be moved from T1 to T3 if it is
already in T1, or just added to T1. However, if the state is in T3 only
mode, a check is done to see if the amount of T1 hits are high enough. The
thought behind this is that if a certain percentage of the hits are in T1
already, we can assume that more of the hits will be there, and we will switch
to tiered mode. Exactly what the percentage of hits should be enough to
switch should be decided experimentally.

In the same way, if the entry is already in T3, a switch from tiered mode to
T3 only mode is performed if the T1 ratio is lower than a set limit.

\subsection{Methodology}

Initially the standard DCPT algorithm was implemented and tested. The prefetcher ran 
with a speedup comparative to the results of Grannæs et. al. As  DTDCPT
is an extension of the DCPT algorithm, this was beneficial to have as a
starting point.

The DTDCPT implementation was done while developing the algorithm, as the logic
for switching tier modes was independent of the code running the tiers. When the
logic was done, it was implemented in the algorithm with preliminary values,
to be changed experimentally.


